Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-27 article.tex
section 4

The top of Table \ref{tab:comparison} reports AUROC and AUPRC for all
four networks using in each case partial correlation with different
filtering functions. Except for the last two rows that use PCA, the
exact inverse of the covariance matrix was used in each case. These
results clearly show the importance of the filters. AUROC increases in
average from 0.77 to 0.93. PCA does not really affect AUROC scores but
it improves significantly AUPRC scores. Taking the average over
various parameter setting gives an improvement of 10\% in AUPRC but
only a minor change in AUROC. The last row shows the final performance
of the method specifically tuned for the challenge (see the Appendix
\ref{app:optimized} for all details). Although this tuning was
decisive to obtain the best performance at the challenge, it does not
significantly improve neither AUROC nor AUPRC.

@@

In the bottom of Table \ref{tab:comparison}, we provide as a comparison the
performance of three other methods: standard (Pearson) correlation (PC),
generalized transfer entropy (GTE), and GENIE3. ROC and PR curves on the
\textit{normal-2} network are shown for all methods in Figure
\ref{fig:curves}. Pearson correlation measures the unconditional linear
(in)dependence between variables and it should thus not be able to filter out
indirect interactions between neurons. GTE \citep{stetter2012model} was
proposed as a baseline for the challenge. This method builds on Transfer
Entropy to measure the association between two neurons. Unlike our approach, it
can predict the direction of the edges. GENIE3 \citep{huynhthu2010inferring} is
a gene regulatory network inference method that was the best performer of the
DREAM5 challenge \citep{marbach2012}. When transposed to neural networks, this
method uses as a confidence score for the edge going from neuron $i$ to neuron
$j$ the importance score of variable $X_i$ in a Random Forest model trying to
predict $X_j$ from all variables in $X\setminus X_j$. To reduce the
computational cost of this method, we had to limit each tree in the Random
Forest model to a maximum depth of 3. This constraint potentially severly
affects the performance of this method with respect to the use of fully grown
trees. We nevertheless provide these results for comparison purpose. PC and
GENIE3 were both applied on the same pre-filtered time series ($g \circ
f_1$). For GENIE3, we built 10000 trees per neuron and we used default settings
for all other parameters (except for the maximal tree depth). For GTE, we
reproduced the exact same setting (conditioning level and pre-processing) that
was used for the challenge.

@@

Partial correlation and average partial correlation clearly outperform all
other methods. The improvement is more important in terms of AUPRC than in
terms of AUROC. As expected, Pearson correlation performs very poorly in terms
of AUPRC. GTE and GENIE3 works much better but these two methods are
nevertheless clearly below partial correlation. Among these two methods, GTE is
slightly better in terms of AUROC, while GENIE3 is significantly better in
terms of AUPRC. Given that we had to limit this latter method for computational
reasons, these results are very promising and a comparison with the full GENIE3
approach is certainly part of our future works.

@@

\subsection*{ROC versus PR curves.}

>>>>

The top of Table \ref{tab:comparison} reports AUROC and AUPRC for all
four networks using in each case partial correlation with different
filtering functions. Except for the last two rows that use PCA, the
exact inverse of the covariance matrix was used in each case. These
results clearly show the importance of the filters. AUROC increases in
average from 0.77 to 0.93. PCA does not really affect AUROC scores but
it improves significantly AUPRC scores. Taking the average over
various parameter setting gives an improvement of 10\% in AUPRC but
only a minor change in AUROC. The last row (``Full method'') shows the
final performance of the method specifically tuned for the challenge
(see the Appendix \ref{app:optimized} for all details). Although this
tuning was decisive to obtain the best performance at the challenge,
it does not significantly improve neither AUROC nor AUPRC.

@@

In the bottom of Table \ref{tab:comparison}, we provide as a comparison the
performance of three other methods: standard (Pearson) correlation (PC),
generalized transfer entropy (GTE), and GENIE3. ROC and PR curves on the
\textit{normal-2} network are shown for all methods in Figure
\ref{fig:curves}. Pearson correlation measures the unconditional linear
(in)dependence between variables and it should thus not be able to filter out
indirect interactions between neurons. GTE \citep{stetter2012model} was
proposed as a baseline for the challenge. This method builds on Transfer
Entropy to measure the association between two neurons. Unlike our approach, it
can predict the direction of the edges. GENIE3 \citep{huynhthu2010inferring} is
a gene regulatory network inference method that was the best performer of the
DREAM5 challenge \citep{marbach2012}. When transposed to neural networks, this
method uses as a confidence score for the edge going from neuron $i$ to neuron
$j$ the importance score of variable $X_i$ in a Random Forest model trying to
predict $X_j$ from all variables in $X\setminus X_j$. To reduce the
computational cost of this method, we had however to limit each tree in the
Random Forest model to a maximum depth of 3. This constraint potentially
severly affects the performance of this method with respect to the use of fully
grown trees. We nevertheless provide these results for comparison purpose. PC
and GENIE3 were both applied on the same pre-filtered time series as partial
correlation ($h\circ g \circ f_1$). For GENIE3, we built 10000 trees per neuron
and we used default settings for all other parameters (except for the maximal
tree depth). For GTE, we reproduced the exact same setting (conditioning level
and pre-processing) that was used for the challenge.

@@

Partial correlation and average partial correlation clearly outperform all
other methods. The improvement is more important in terms of AUPRC than in
terms of AUROC. As expected, Pearson correlation performs very poorly in terms
of AUPRC. GTE and GENIE3 work much better but these two methods are
nevertheless clearly below partial correlation. Among these two methods, GTE is
slightly better in terms of AUROC, while GENIE3 is significantly better in
terms of AUPRC. Given that we had to limit this latter method for computational
reasons, these results are very promising and a comparison with the full GENIE3
approach is certainly part of our future works.

The fact that our method is unable to predict edge directions does not seem to
be a disadvantage with respect to GTE and GENIE3. Although partial correlation
scores each edge and its opposite similarly, it can reach precision values
higher than 0.5 (see Figure \ref{fig:curves}(b)), suggesting that it ranks high
mainly pairs of neurons that interact in both directions (there are about XX\%
of such pairs in the \textit{normal-2} network). %% It is interesting also to note
%% that, on \textit{normal-2}, a method that perfectly predicts the undirected
%% network (i.e. that gives a score of 1 to each pair $(i,j)$ such that $(i,j)\in
%% V$ or $(j,i)\in V$, 0 otherwise) can reach an AUROC of $0.995$, which does not
%% leave much room for improvement by predicting edge directionality. The AUPRC on
%% the other hand would be 0.789 and thus be more sensitive to the prediction of
%% edge directions.

@@

%\subsection*{ROC versus PR curves.}
